---
title: "Riddler Submission 2/10/2020"
author: "Tomas Okal"
date: "2/10/2020"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(stringi)
library(knitr)
```

# Riddle

"From the James Anderson comes a palindromic puzzle of calendars:

This past Sunday was Groundhog Day. Also, there was a football game. But to top it all off, the date, 02/02/2020, was palindromic, meaning it reads the same forwards and backwards (if you ignore the slashes).

If we write out dates in the American format of MM/DD/YYYY (i.e., the two digits of the month, followed by the two digits of the day, followed by the four digits of the year), how many more palindromic dates will there be this century?"

# Parameters

Assuming that the last day of this century is December 31st, 2099, then we can construct a sequence of dates from 02-02-2020 to 12-31-2099 using the `sewq` and `as.Date` function in R. This will let us specify the unit of time as well.

```{r}
head(seq(as.Date("2020-02-02"), as.Date("2099-12-31"), by = "days"))
tail(seq(as.Date("2020-02-02"), as.Date("2099-12-31"), by = "days"))
length(seq(as.Date("2020-02-02"), as.Date("2099-12-31"), by = "days"))
```

From this, we can see that we have a total of 29,188 days left this century.

```{r possible_dates}
possible_days <- seq(as.Date("2020-02-02"), as.Date("2099-12-31"), by = "days")

df_dates <- data.frame(possible_days)
df_dates$y_chr <- as.numeric(format(as.Date(possible_days), "%Y"))
df_dates$m_chr <- as.character(format.Date(possible_days, "%m"))
df_dates$d_chr <- as.character(format.Date(possible_days, "%d"))
```

Using the `as.Date` function, we can extract the year, month, and day from this sequence.

```{r, echo = FALSE}
knitr::kable(tail(df_dates), format = "markdown")
```

Now that we have the individual elements needed, we can paste together the month and day compenents with `paste0`. After this, the `stringi` package lets us reverse the resulting character vector with `stri_reverse`. I convert this character vector output to a numeric so that I can match it.

```{r}
df_dates$md_chr <- paste0(df_dates$m_chr, df_dates$d_chr)
df_dates$md_rev <- as.numeric(stringi::stri_reverse(df_dates$md_chr))
```

Now we can use a simple `ifelse` to match the original year and created reverse of month + day. All of the matches are then put in the new object matches.

```{r}
df_dates$match <- ifelse(df_dates$y_chr == df_dates$md_rev, 1, 0)
matches <- df_dates[df_dates$match == 1, ]
```

Filtering to only the dates that have a palindromic match, we get:

```{r}
knitr::kable(matches, format = "markdown")
```


# Answer

From the 29188 days left in this century, we have 9 more palindromic dates! 

